probKanren: A Simple Probabilistic extension for microKanren

### 2021-09-14 paper/main.tex
Add semantics paragraph

@@

\textsc{microKanren}\cite{10.1145/2989225.2989230, daniel2018reasoned}
is a pure logic programming language embedded in Scheme. The language
consists of a set of terms, a set of goal primitives, and two run
functions to answer queries in the language. The goal primitives
consist of a \texttt{fresh} form for introducing logic variables, a
unification primitive \texttt{==}, a conjunction combinator \texttt{conj}, a disjunction
combinator \texttt{disj}, and ways to define and apply relations.

@@

\begin{figure}
\begin{grammar}
<prog> ::= (run <number> (<id>) <goal-expr>)

@@

\end{grammar}
\end{figure}

@@

Sequential Monte Carlo\cite{chopin2020introduction} are an efficient
online way to sample from probabilistic models especially suited for
state-space domains. If we imagine our probabilistic programs as
straight-line programs with no control-flow we can imagine numbering
every sample function $f_1, f_2, \ldots, f_n$ and every observe
function $g_1, g_2, \ldots, g_n$ then our probability density over our
random variables $x$ and observed data $y$ can be defined as:

@@

We propose to extend \textsc{microKanren} by augmenting each of the
search streams with a set of particles. These particles represent the
empirical distribution of that stream. Each particle has associated
with it a substitution of all the logic and random variables as well
as a weight that is proportional to the likelihood of the substitution.

@@

>>>>

\usepackage{cleveref}

@@

\textsc{microKanren}\cite{10.1145/2989225.2989230, daniel2018reasoned}
is a pure logic programming language embedded in Scheme. The language
consists of a set of terms, a set of goal primitives, and two run
functions to answer queries in the language. The goal primitives
consist of a \texttt{fresh} form for introducing logic variables, a
unification primitive \texttt{==}, a conjunction combinator
\texttt{conj}, a disjunction combinator \texttt{disj}, and ways to
define and apply relations. Further forms are shown in detail in
\Cref{fig:grammar}.

@@

\begin{figure}
  \centering
  \begin{grammar}
<prog> ::= (run <number> (<id>) <goal-expr>)

@@

  \end{grammar}
  \caption{Grammar for \textsc{microKanren}}
  \label{fig:grammar}
\end{figure}

@@

Sequential Monte Carlo\cite{chopin2020introduction}(SMC) is an efficient
online way to sample from probabilistic models especially suited for
state-space domains. If we imagine our probabilistic programs as
straight-line programs with no control-flow we can imagine numbering
every sample function $f_1, f_2, \ldots, f_n$ and every observe
function $g_1, g_2, \ldots, g_n$ then our probability density over our
random variables $x$ and observed data $y$ can be defined as:

@@

We propose to extend \textsc{microKanren} by turning the search into a
SMC sampler. We accomplish this by augmenting each of the search
streams with a set of particles. These particles represent the
empirical distribution of that stream. Each particle has associated
with it a substitution of all the logic and random variables as well
as a weight that is proportional to the likelihood of the
substitution.

@@

We follow the semantics of \cite{staton2016semantics}, where each
language form has a denotational semantics that associates a
measurable function with each form, and an operational semantics that associates
a sampler with each form.
